### 2023-12-09 paper/paper.tex

%
% File acl2019.tex
%
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2019}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amssymb}

\usepackage{url}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{Instructions for ACL 2019 Proceedings}

\author{George Vallejo \\
  \texttt{gavallejo02@uchicago.edu} \\\And
  Nick Clifford \\
  \texttt{nclifford@uchicago.edu} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  Several features of human language can be described through rules,
  and natively speaking children of languages are cognizant of these rules, 
  even without formal instruction. Prior research has attempted to model
  such learning in children using connectionist neural networks for
  paradigms such as the English past tense. We investigate how a gradient-descent
  neural network could acquire the rules for declining Russian nouns into
  the genitive case for both singular and plural nouns in a similar fashion. 
  We also test whether two different NNs for the singular and plural numbers
  can model the learning of these rules in children better than a single
  NN by comparing the models' performance to that of native Russian-speaking children.

\end{abstract}

\section{Credits}

  This paper is an adaptation of the paper
  \emph{On Learning the Past Tenses of English Verbs} by D. E. Rumelhart and
  J. L. McClelland.

\section{Introduction}

Many features of natural languages, such as verb conjugation, noun case
declension, and agglutination behave in ways that can be described
by sets of rules. Being cognizant of these rules is an important aspect
of language acquisition in children, because awareness of these rules allows
them to utilize newly acquired vocabulary to express new concepts in the same
manner they may have used in a previous utterance. For example, the most common
ending used to express the past tense in English is -ed, as in ``I waited at home''.
Awareness of this rule would allow a child to additionally express ``I hunted at
the park'' after learning the verb ``hunt'' using the same ending.

However, it is also often the case that these rules have exceptions or other
irregularities. In English, ablaut is a source of several irregular verbs'
past tense forms. For example, the past tense of ``run'' is ``ran'', and
a child aware of the basic past tense rule may incorrectly express
``I runned to the park'' as a result, a phenomenon known as overgeneralizing.
This phemoenon is also important for learning, as the corrections the child
receives improve their understanding of the past tense, ultimately allowing
them to generalize the past tense for more verbs with greater accuracy.

Prior research has attempted to create neural networks to function as
language acquisition devices (LAD) in order to model how children acquire
the English past tense. Most notably, Rumelhart and McClelland (1986)
made use of a perceptron model to model the learning of English past
tense in three stages based on prior research into how children
acquire it, and obtained learning results that very closely
align with children's learning. We therefore investigate whether
the acquisition of the genitive case in Russian would yield similar results.

\subsection{Russian noun case morphology}

Grammatical cases are a system of marking nouns and noun modifiers in order
to indicate the noun (modifier)'s intended grammatical function in a sentence.
While historically English had grammatical cases, they survive in modern English
only in pronouns. Every pronoun in English has three different forms representing
three different cases. These are:

\begin{enumerate}
    \item Nominative: Marks the subject of a sentence: \emph{I} help him.
    \item Accusative: Marks the indirect object of a sentence: He helps \emph{me}.
    \item Genitive: Marks the possessor of a noun: \emph{My} book is great.
\end{enumerate}

Russian, being a member of the Slavic branch of the Indo-European languages,
has a fully functioning grammatical case system applying for all nouns
unlike English. Nouns and noun modifiers in Russian inflect for case
by changing the ending of the word, and the exact manner by which an ending
changes depends on the ending, case, and number of the noun.

Table 1 shows the different endings of the Russian nominative and
genitive cases in the singular and plural.

\begin{table}[t!]
\begin{center}
\begin{tabular}{|l|l|l}
\hline & \textbf{Singular} & \textbf{Plural} \\ \hline
\textbf{Nominative} & -a, -ja, -ija & -y, -i, -ii \\
\textbf{Genitive} & -y, -i, -ii & $\varnothing$, -', -ij \\ \hline
\textbf{Nominative} & $\varnothing$, -'/j, -ij & -y, -i, -ii, -je \\
\textbf{Genitive} & -a, -ja, -ija & -ov, -jej/jev, -ijev \\ \hline
\textbf{Nominative} & -o, -je & -a, -ja \\
\textbf{Genitive} & -a, -ja & $\varnothing$, -j, -jej \\ \hline
\textbf{Nominative} & -' & -i \\
\textbf{Genitive} & -i & -jej \\ \hline
\textbf{Nominative} & -mja & -mjena \\
\textbf{Genitive} & -mjeni & -mjon \\ \hline
\end{tabular}
\end{center}
\caption{\label{declension} Nominative and genitive endings for nouns with different endings. Note that the apostrophe ' is a palatalization character, corresponding to \foreignlanguage{russian}{ь} in Cyrillic. }
\end{table}

\section{General Instructions}

Manuscripts must be in two-column format.  Exceptions to the
two-column format include the title, authors' names and complete
addresses, which must be centered at the top of the first page, and
any full-width figures or tables (see the guidelines in
Subsection~\ref{ssec:first}). \textbf{Type single-spaced.}  Start all
pages directly under the top margin. See the guidelines later
regarding formatting the first page.  The manuscript should be
printed single-sided and its length
should not exceed the maximum page limit described in Section~\ref{sec:length}.
Pages are numbered for  initial submission. However, \textbf{do not number the pages in the camera-ready version}.

By uncommenting {\small\verb|\aclfinalcopy|} at the top of this 
 document, it will compile to produce an example of the camera-ready formatting; by leaving it commented out, the document will be anonymized for initial submission.  When you first create your submission on softconf, please fill in your submitted paper ID where {\small\verb|***|} appears in the {\small\verb|\def\aclpaperid{***}|} definition at the top.

The review process is double-blind, so do not include any author information (names, addresses) when submitting a paper for review.  
However, you should maintain space for names and addresses so that they will fit in the final (accepted) version.  The ACL 2019 \LaTeX\ style will create a titlebox space of 2.5in for you when {\small\verb|\aclfinalcopy|} is commented out.

\subsection{The Ruler}
The ACL 2019 style defines a printed ruler which should be presented in the
version submitted for review.  The ruler is provided in order that
reviewers may comment on particular lines in the paper without
circumlocution.  If you are preparing a document without the provided
style files, please arrange for an equivalent ruler to
appear on the final output pages.  The presence or absence of the ruler
should not change the appearance of any other content on the page.  The
camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment the {\small\verb|\aclfinalcopy|} command in the document preamble.)

Reviewers: note that the ruler measurements do not align well with
lines in the paper -- this turns out to be very difficult to do well
when the paper contains many figures and equations, and, when done,
looks ugly. In most cases one would expect that the approximate
location will be adequate, although you can also use fractional
references (\emph{e.g.}, the first paragraph on this page ends at mark $108.5$).

\subsection{Electronically-available resources}

ACL provides this description in \LaTeX2e{} (\texttt{\small acl2019.tex}) and PDF
format (\texttt{\small acl2019.pdf}), along with the \LaTeX2e{} style file used to
format it (\texttt{\small acl2019.sty}) and an ACL bibliography style (\texttt{\small acl\_natbib.bst})
and example bibliography (\texttt{\small acl2019.bib}).
These files are all available at
\texttt{\small http://acl2019.org/downloads/ acl2019-latex.zip}. 
 We
strongly recommend the use of these style files, which have been
appropriately tailored for the ACL 2019 proceedings.

\subsection{Format of Electronic Manuscript}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's
Portable Document Format (PDF). PDF files are usually produced from
\LaTeX\ using the \textit{pdflatex} command. If your version of
\LaTeX\ produces Postscript files, you can convert these into PDF
using \textit{ps2pdf} or \textit{dvipdf}. On Windows, you can also use
Adobe Distiller to generate PDF.

Please make sure that your PDF file includes all the necessary fonts
(especially tree diagrams, symbols, and fonts with Asian
characters). When you print or create the PDF file, there is usually
an option in your printer setup to include none, all or just
non-standard fonts.  Please make sure that you select the option of
including ALL the fonts. \textbf{Before sending it, test your PDF by
  printing it from a computer different from the one where it was
  created.} Moreover, some word processors may generate very large PDF
files, where each page is rendered as an image. Such images may
reproduce poorly. In this case, try alternative ways to obtain the
PDF. One way on some systems is to install a driver for a postscript
printer, send your document to the printer specifying ``Output to a
file'', then convert the file to PDF.

It is of utmost importance to specify the \textbf{A4 format} (21 cm
x 29.7 cm) when formatting the paper. When working with
\texttt{dvips}, for instance, one should specify \texttt{-t a4}.
Or using the command \verb|\special{papersize=210mm,297mm}| in the latex
preamble (directly below the \verb|\usepackage| commands). Then using 
\texttt{dvipdf} and/or \texttt{pdflatex} which would make it easier for some.

Print-outs of the PDF file on A4 paper should be identical to the
hardcopy version. If you cannot meet the above requirements about the
production of your electronic submission, please contact the
publication chairs as soon as possible.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts two columns to a page, in the manner these
instructions are formatted. The exact dimensions for a page on A4
paper are:

\begin{itemize}
\item Left and right margins: 2.5 cm
\item Top margin: 2.5 cm
\item Bottom margin: 2.5 cm
\item Column width: 7.7 cm
\item Column height: 24.7 cm
\item Gap between columns: 0.6 cm
\end{itemize}

\noindent Papers should not be submitted on any other paper size.
 If you cannot meet the above requirements about the production of 
 your electronic submission, please contact the publication chairs 
 above as soon as possible.

\subsection{Fonts}

For reasons of uniformity, Adobe's \textbf{Times Roman} font should be
used. In \LaTeX2e{} this is accomplished by putting

\begin{quote}
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}
in the preamble. If Times Roman is unavailable, use \textbf{Computer
  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
  10\% less dense than Adobe's Times Roman font.

\begin{table}[t!]
\begin{center}
\begin{tabular}{|l|rl|}
\hline \textbf{Type of Text} & \textbf{Font Size} & \textbf{Style} \\ \hline
paper title & 15 pt & bold \\
author names & 12 pt & bold \\
author affiliation & 12 pt & \\
the word ``Abstract'' & 12 pt & bold \\
section titles & 12 pt & bold \\
subsection titles & 11 pt & bold \\
document text & 11 pt  &\\
captions & 10 pt & \\
abstract text & 10 pt & \\
bibliography & 10 pt & \\
footnotes & 9 pt & \\
\hline
\end{tabular}
\end{center}
\caption{\label{font-table} Font guide. }
\end{table}

\subsection{The First Page}
\label{ssec:first}

Center the title, author's name(s) and affiliation(s) across both
columns. Do not use footnotes for affiliations. Do not include the
paper ID number assigned during the submission process. Use the
two-column format only when you begin the abstract.

\textbf{Title}: Place the title centered at the top of the first page, in
a 15-point bold font. (For a complete guide to font sizes and styles,
see Table~\ref{font-table}) Long titles should be typed on two lines
without a blank line intervening. Approximately, put the title at 2.5
cm from the top of the page, followed by a blank line, then the
author's names(s), and the affiliation on the following line. Do not
use only initials for given names (middle initials are allowed). Do
not format surnames in all capitals (\emph{e.g.}, use ``Mitchell'' not
``MITCHELL'').  Do not format title and section headings in all
capitals as well except for proper names (such as ``BLEU'') that are
conventionally in all capitals.  The affiliation should contain the
author's complete address, and if possible, an electronic mail
address. Start the body of the first page 7.5 cm from the top of the
page.

The title, author names and addresses should be completely identical
to those entered to the electronical paper submission website in order
to maintain the consistency of author information among all
publications of the conference. If they are different, the publication
chairs may resolve the difference without consulting with you; so it
is in your own interest to double-check that the information is
consistent.

\textbf{Abstract}: Type the abstract at the beginning of the first
column. The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by about
0.6 cm on each side. Center the word \textbf{Abstract} in a 12 point bold
font above the body of the abstract. The abstract should be a concise
summary of the general thesis and conclusions of the paper. It should
be no longer than 200 words. The abstract text should be in 10 point font.

\textbf{Text}: Begin typing the main body of the text immediately after
the abstract, observing the two-column format as shown in the present document. Do not include page numbers.

\textbf{Indent}: Indent when starting a new paragraph, about 0.4 cm. Use 11 points for text and subsection headings, 12 points for section headings and 15 points for the title.

\begin{table}
\centering
\small
\begin{tabular}{cc}
\begin{tabular}{|l|l|}
\hline
\textbf{Command} & \textbf{Output}\\\hline
\verb|{\"a}| & {\"a} \\
\verb|{\^e}| & {\^e} \\
\verb|{\`i}| & {\`i} \\ 
\verb|{\.I}| & {\.I} \\ 
\verb|{\o}| & {\o} \\
\verb|{\'u}| & {\'u}  \\ 
\verb|{\aa}| & {\aa}  \\\hline
\end{tabular} & 
\begin{tabular}{|l|l|}
\hline
\textbf{Command} & \textbf{Output}\\\hline
\verb|{\c c}| & {\c c} \\ 
\verb|{\u g}| & {\u g} \\ 
\verb|{\l}| & {\l} \\ 
\verb|{\~n}| & {\~n} \\ 
\verb|{\H o}| & {\H o} \\ 
\verb|{\v r}| & {\v r} \\ 
\verb|{\ss}| & {\ss} \\\hline
\end{tabular}
\end{tabular}
\caption{Example commands for accented characters, to be used in, \emph{e.g.}, \BibTeX\ names.}\label{tab:accents}
\end{table}

\subsection{Sections}

\textbf{Headings}: Type and label section and subsection headings in the
style shown on the present document.  Use numbered sections (Arabic
numerals) in order to facilitate cross references. Number subsections
with the section number and the subsection number separated by a dot,
in Arabic numerals.
Do not number subsubsections.

\begin{table*}[t!]
\centering
\begin{tabular}{lll}
  output & natbib & previous ACL style files\\
  \hline
  \citep{Gusfield:97} & \verb|\citep| & \verb|\cite| \\
  \citet{Gusfield:97} & \verb|\citet| & \verb|\newcite| \\
  \citeyearpar{Gusfield:97} & \verb|\citeyearpar| & \verb|\shortcite| \\
\end{tabular}
\caption{Citation commands supported by the style file.
  The citation style is based on the natbib package and
  supports all natbib citation commands.
  It also supports commands defined in previous ACL style files
  for compatibility.
  }
\end{table*}

\textbf{Citations}: Citations within the text appear in parentheses
as~\cite{Gusfield:97} or, if the author's name appears in the text
itself, as Gusfield~\shortcite{Gusfield:97}.
Using the provided \LaTeX\ style, the former is accomplished using
{\small\verb|\cite|} and the latter with {\small\verb|\shortcite|} or {\small\verb|\newcite|}. Collapse multiple citations as in~\cite{Gusfield:97,Aho:72}; this is accomplished with the provided style using commas within the {\small\verb|\cite|} command, \emph{e.g.}, {\small\verb|\cite{Gusfield:97,Aho:72}|}. Append lowercase letters to the year in cases of ambiguities.  
 Treat double authors as
in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two
authors are involved. Collapse multiple citations as
in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations
as sentence constituents.

We suggest that instead of
\begin{quote}
  ``\cite{Gusfield:97} showed that ...''
\end{quote}
you use
\begin{quote}
``Gusfield \shortcite{Gusfield:97}   showed that ...''
\end{quote}

If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
can use the command \verb|\citet| (cite in text)
to get ``author (year)'' citations.

You can use the command \verb|\citealp| (alternative cite without 
parentheses) to get ``author year'' citations (which is useful for 
using citations within parentheses, as in \citealp{Gusfield:97}).

If the Bib\TeX{} file contains DOI fields, the paper
title in the references section will appear as a hyperlink
to the DOI, using the hyperref \LaTeX{} package.
To disable the hyperref package, load the style file
with the \verb|nohyperref| option: \\{\small
\verb|\usepackage[nohyperref]{acl2019}|}

\textbf{Compilation Issues}: Some of you might encounter the following error during compilation:

``{\em \verb|\pdfendlink| ended up in different nesting level than \verb|\pdfstartlink|.}''

This happens when \verb|pdflatex| is used and a citation splits across a page boundary. To fix this, the style file contains a patch consisting of the following two lines: (1) \verb|\RequirePackage{etoolbox}| (line 454 in \texttt{acl2019.sty}), and (2) A long line below (line 455 in \texttt{acl2019.sty}).

If you still encounter compilation issues even with the patch enabled, disable the patch by commenting the two lines, and then disable the \verb|hyperref| package (see above), recompile and see the problematic citation.
Next rewrite that sentence containing the citation. (See, {\em e.g.}, {\small\tt http://tug.org/errors.html})

\textbf{Digital Object Identifiers}:  As part of our work to make ACL
materials more widely used and cited outside of our discipline, ACL
has registered as a CrossRef member, as a registrant of Digital Object
Identifiers (DOIs), the standard for registering permanent URNs for
referencing scholarly materials.  As of 2017, we are requiring all
camera-ready references to contain the appropriate DOIs (or as a
second resort, the hyperlinked ACL Anthology Identifier) to all cited
works.  Thus, please ensure that you use Bib\TeX\ records that contain
DOI or URLs for any of the ACL materials that you reference.
Appropriate records should be found for most materials in the current
ACL Anthology at \url{http://aclanthology.info/}.

As examples, we cite \cite{P16-1001} to show you how papers with a DOI
will appear in the bibliography.  We cite \cite{C14-1001} to show how
papers without a DOI but with an ACL Anthology Identifier will appear
in the bibliography.

As reviewing will be double-blind, the submitted version of the papers
should not include the authors' names and affiliations. Furthermore,
self-references that reveal the author's identity, \emph{e.g.},
\begin{quote}
``We previously showed \cite{Gusfield:97} ...''  
\end{quote}
should be avoided. Instead, use citations such as 
\begin{quote}
``\citeauthor{Gusfield:97} \shortcite{Gusfield:97}
previously showed ... ''
\end{quote}

Any preliminary non-archival versions of submitted papers should be listed in the submission form but not in the review version of the paper. ACL 2019 reviewers are generally aware that authors may present preliminary versions of their work in other venues, but will not be provided the list of previous presentations from the submission form.

\textbf{Please do not use anonymous citations} and do not include
 when submitting your papers. Papers that do not
conform to these requirements may be rejected without review.

\textbf{References}: Gather the full set of references together under
the heading \textbf{References}; place the section before any Appendices. 
Arrange the references alphabetically
by first author, rather than by order of occurrence in the text.
By using a .bib file, as in this template, this will be automatically 
handled for you. See the \verb|\bibliography| commands near the end for more.

Provide as complete a citation as possible, using a consistent format,
such as the one for \emph{Computational Linguistics\/} or the one in the 
\emph{Publication Manual of the American 
Psychological Association\/}~\cite{APA:83}. Use of full names for
authors rather than initials is preferred. A list of abbreviations
for common computer science journals can be found in the ACM 
\emph{Computing Reviews\/}~\cite{ACM:83}.

The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
American Psychological Association format, allowing regular citations, 
short citations and multiple citations as described above.

\begin{itemize}
\item Example citing an arxiv paper: \cite{rasooli-tetrault-2015}. 
\item Example article in journal citation: \cite{Ando2005}.
\item Example article in proceedings, with location: \cite{borsch2011}.
\item Example article in proceedings, without location: \cite{andrew2007scalable}.
\end{itemize}
See corresponding .bib file for further details.

Submissions should accurately reference prior and related work, including code and data. If a piece of prior work appeared in multiple venues, the version that appeared in a refereed, archival venue should be referenced. If multiple versions of a piece of prior work exist, the one used by the authors should be referenced. Authors should not rely on automated citation indices to provide accurate references for prior and related work.

\textbf{Appendices}: Appendices, if any, directly follow the text and the
references (but see above).  Letter them in sequence and provide an
informative title: \textbf{Appendix A. Title of Appendix}.

\subsection{Footnotes}

\textbf{Footnotes}: Put footnotes at the bottom of the page and use 9
point font. They may be numbered or referred to by asterisks or other
symbols.\footnote{This is how a footnote should appear.} Footnotes
should be separated from the text by a line.\footnote{Note the line
separating the footnotes from the text.}

\subsection{Graphics}

\textbf{Illustrations}: Place figures, tables, and photographs in the
paper near where they are first discussed, rather than at the end, if
possible.  Wide illustrations may run across both columns.  Color
illustrations are discouraged, unless you have verified that  
they will be understandable when printed in black ink.

\textbf{Captions}: Provide a caption for every illustration; number each one
sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
Caption of the Table.''  Type the captions of the figures and 
tables below the body, using 10 point text. Captions should be placed below illustrations. Captions that are one line are centered (see Table~\ref{font-table}). Captions longer than one line are left-aligned (see Table~\ref{tab:accents}). Do not overwrite the default caption sizes. The acl2019.sty file is compatible with the caption and subcaption packages; do not add optional arguments.

\subsection{Accessibility}
\label{ssec:accessibility}

In an effort to accommodate people who are color-blind (as well as those printing
to paper), grayscale readability for all accepted papers will be
encouraged.  Color is not forbidden, but authors should ensure that
tables and figures do not rely solely on color to convey critical
distinctions. A simple criterion: All curves and points in your figures should be clearly distinguishable without color.

% Min: no longer used as of ACL 2018, following ACL exec's decision to
% remove this extra workflow that was not executed much.
% BEGIN: remove
%% \section{XML conversion and supported \LaTeX\ packages}

%% Following ACL 2014 we will also we will attempt to automatically convert 
%% your \LaTeX\ source files to publish papers in machine-readable 
%% XML with semantic markup in the ACL Anthology, in addition to the 
%% traditional PDF format.  This will allow us to create, over the next 
%% few years, a growing corpus of scientific text for our own future research, 
%% and picks up on recent initiatives on converting ACL papers from earlier 
%% years to XML.

%% We encourage you to submit a ZIP file of your \LaTeX\ sources along
%% with the camera-ready version of your paper. We will then convert them
%% to XML automatically, using the LaTeXML tool
%% (\url{http://dlmf.nist.gov/LaTeXML}). LaTeXML has \emph{bindings} for
%% a number of \LaTeX\ packages, including the ACL 2018 stylefile. These
%% bindings allow LaTeXML to render the commands from these packages
%% correctly in XML. For best results, we encourage you to use the
%% packages that are officially supported by LaTeXML, listed at
%% \url{http://dlmf.nist.gov/LaTeXML/manual/included.bindings}
% END: remove

\section{Translation of non-English Terms}

It is also advised to supplement non-English characters and terms
with appropriate transliterations and/or translations
since not all readers understand all such characters and terms.
Inline transliteration or translation can be represented in
the order of: original-form transliteration ``translation''.

\section{Length of Submission}
\label{sec:length}

The ACL 2019 main conference accepts submissions of long papers and
short papers.
 Long papers may consist of up to eight (8) pages of
content plus unlimited pages for references. Upon acceptance, final
versions of long papers will be given one additional page -- up to nine (9)
pages of content plus unlimited pages for references -- so that reviewers' comments
can be taken into account. Short papers may consist of up to four (4)
pages of content, plus unlimited pages for references. Upon
acceptance, short papers will be given five (5) pages in the
proceedings and unlimited pages for references. 
For both long and short papers, all illustrations and tables that are part
of the main text must be accommodated within these page limits, observing
the formatting instructions given in the present document. Papers that do not conform to the specified length and formatting requirements are subject to be rejected without review.

ACL 2019 does encourage the submission of additional material that is relevant to the reviewers but not an integral part of the paper. There are two such types of material: appendices, which can be read, and non-readable supplementary materials, often data or code.  Do not include this additional material in the same document as your main paper. Additional material must be submitted as one or more separate files, and must adhere to the same anonymity guidelines as the main paper. The paper must be self-contained: it is optional for reviewers to look at the supplementary material. Papers should not refer, for further detail, to documents, code or data resources that are not available to the reviewers. Refer to Appendix~\ref{sec:appendix} and Appendix~\ref{sec:supplemental} for further information.

Workshop chairs may have different rules for allowed length and
whether supplemental material is welcome. As always, the respective
call for papers is the authoritative source.

\section*{Acknowledgments}

The acknowledgments should go immediately before the references.  Do
not number the acknowledgments section. Do not include this section
when submitting your paper for review. \\

\noindent \textbf{Preparing References:} \\
Include your own bib file like this:
\verb|\bibliographystyle{acl_natbib}|
\verb|\bibliography{acl2019}|

where \verb|acl2019| corresponds to a acl2019.bib file.
\bibliography{acl2019}
\bibliographystyle{acl_natbib}

\appendix

\section{Appendices}
\label{sec:appendix}
Appendices are material that can be read, and include lemmas, formulas, proofs, and tables that are not critical to the reading and understanding of the paper. 
Appendices should be \textbf{uploaded as supplementary material} when submitting the paper for review. Upon acceptance, the appendices come after the references, as shown here. Use
\verb|\appendix| before any appendix section to switch the section
numbering over to letters.

\section{Supplemental Material}
\label{sec:supplemental}
Submissions may include non-readable supplementary material used in the work and described in the paper. Any accompanying software and/or data should include licenses and documentation of research review as appropriate. Supplementary material may report preprocessing decisions, model parameters, and other details necessary for the replication of the experiments reported in the paper. Seemingly small preprocessing decisions can sometimes make a large difference in performance, so it is crucial to record such decisions to precisely characterize state-of-the-art methods.

Nonetheless, supplementary material should be supplementary (rather
than central) to the paper. \textbf{Submissions that misuse the supplementary 
material may be rejected without review.}
Supplementary material may include explanations or details
of proofs or derivations that do not fit into the paper, lists of
features or feature templates, sample inputs and outputs for a system,
pseudo-code or source code, and data. (Source code and data should
be separate uploads, rather than part of the paper).

The paper should not rely on the supplementary material: while the paper
may refer to and cite the supplementary material and the supplementary material will be available to the
reviewers, they will not be asked to review the
supplementary material.

\end{document}

### 2023-12-09 paper/paper.tex
further paper contribution

@@
 \usepackage{amssymb}
+\usepackage{multirow}
+\usepackage{tipa}
 
@@
 While historically English had grammatical cases, they survive in modern English
-only in pronouns. Every pronoun in English has three different forms representing
+only through pronouns. Every pronoun in English has three different forms representing
 three different cases. These are:
@@
 Russian, being a member of the Slavic branch of the Indo-European languages,
-has a fully functioning grammatical case system applying for all nouns
+has a fully functioning grammatical case system which is used for all nouns
 unlike English. Nouns and noun modifiers in Russian inflect for case
 by changing the ending of the word, and the exact manner by which an ending
-changes depends on the ending, case, and number of the noun.
+changes depends on the ending, case, and number of the noun. Table 1 shows 
+the different possible endings of the Russian nominative and genitive cases in 
+the singular and plural.
 
-Table 1 shows the different endings of the Russian nominative and
-genitive cases in the singular and plural.
+There are, of course, various irregularities in forming the genitive case
+for certain nouns. Some nouns have suppletive roots, found only in the
+plural form: for example, the noun \emph{\v{c}elovjek} (person) has the regular singular 
+genitive form \emph{\v{c}elovjeka}. However, in the nominative plural, it becomes
+\emph{ljudji}, and in the genitive plural \emph{ljudjej}, using an entirely different
+noun stem for the plural cases.
+
+The reason we chose the genitive case to study is because its formation,
+much like the English past tense, is fairly regular in the singular; in a majority of cases,
+the genitive form can be simply derived from the nominative by modifying the
+noun's ending based on the table. We also chose it because the genitive plural,
+in particular, tends to be more irregular. It is often the aspect of Russian
+that is most difficult for learners of the language to comprehend, so
+we wanted to investigate how a gradient-descent neural network would handle
+acquiring the genitive case in singular and plural forms.
+
+We also chose the genitive because it is a case that (with the exception of
+indeclinable nouns, which we ignored for this paper) always changes the noun
+in some way. This is contrasted with the accusative case, which can either
+not change the noun, use the genitive form to represent the accusative, or 
+use its own unique endings. Whichever scenario applies depends on not just 
+the ending of the noun, but also its gender and animacy.
 
@@
 \hline & \textbf{Singular} & \textbf{Plural} \\ \hline
-\textbf{Nominative} & -a, -ja, -ija & -y, -i, -ii \\
-\textbf{Genitive} & -y, -i, -ii & $\varnothing$, -', -ij \\ \hline
-\textbf{Nominative} & $\varnothing$, -'/j, -ij & -y, -i, -ii, -je \\
+\textbf{Nominative} & -a, -ja, -ija & -I\textbackslash, -i, -ii \\
+\textbf{Genitive} & -I\textbackslash, -i, -ii & $\varnothing$, -', -ij \\ \hline
+\textbf{Nominative} & $\varnothing$, -'/j, -ij & -I\textbackslash, -i, -ii, -je \\
 \textbf{Genitive} & -a, -ja, -ija & -ov, -jej/jev, -ijev \\ \hline
@@
 \end{center}
-\caption{\label{declension} Nominative and genitive endings for nouns with different endings.
+\caption{\label{declension} Nominative and genitive endings for nouns with different endings, written with X-SAMPA.
 Note that the apostrophe ' is a palatalization character, corresponding to \foreignlanguage{russian}{ь} in Cyrillic.
@@
 
-\section{General Instructions}
+\section{Featural representation of Russian phonemes}
 
-Manuscripts must be in two-column format.  Exceptions to the
-two-column format include the title, authors' names and complete
-addresses, which must be centered at the top of the first page, and
-any full-width figures or tables (see the guidelines in
-Subsection~\ref{ssec:first}). \textbf{Type single-spaced.}  Start all
-pages directly under the top margin. See the guidelines later
-regarding formatting the first page.  The manuscript should be
-printed single-sided and its length
-should not exceed the maximum page limit described in Section~\ref{sec:length}.
-Pages are numbered for  initial submission. However, \textbf{do not number the pages in the camera-ready version}.
+Rumelhart and McClelland (1986) made use of a scheme to encode
+English phoneme units, known as Wickelphones and Wickelfeatures, described
+in Wickelgren (1969). Wickelphones are phoneme units representing
+each phone in a word as triples, which consist of the phoneme itself,
+its predecessor to the left, and its successor on the right.
+Wickelfeatures are a means of representing Wickelphones as distributed
+patterns of activation by capturing one feature of the central phoneme,
+one feature of the predecessor, and one feature of the successor.
 
-By uncommenting {\small\verb|\aclfinalcopy|} at the top of this 
- document, it will compile to produce an example of the camera-ready formatting; by leaving it commented out, the document will be anonymized for initial submission.
-When you first create your submission on softconf, please fill in your submitted paper ID where {\small\verb|***|} appears in the {\small\verb|\def\aclpaperid{***}|} definition at the top.
+The feature coding scheme we used to capture Russian phonemes did not
+differ heavily from Rumelhart and McClelland's scheme. However, in order
+to accomodate Russian's richer consonant inventory, we had to add an
+extra dimension to the scheme to account for plain vs. palatalized
+consonants, the latter of which are in abundance in Russian phonology.
 
-The review process is double-blind, so do not include any author information (names, addresses) when submitting a paper for review.  
-However, you should maintain space for names and addresses so that they will fit in the final (accepted) version.
-The ACL 2019 \LaTeX\ style will create a titlebox space of 2.5in for you when {\small\verb|\aclfinalcopy|} is commented out.
+In addition to accomodating Russian phonology for Wickelfeatures, we had to make some
+decisions regards to what exact sounds are in Russian. Specifically,
+the status of the close central unrounded vowel \textipa{/\textbari/} 
+(written in Cyrillic as \foreignlanguage{russian}{ы}) as a separate 
+vowel morpheme in Russian is a subject of heavy debate. Some linguists 
+believe in a five-vowel analysis, meaning that the vowel \textipa{/\textbari/} 
+is in complementary distribution with the close front unrounded vowel 
+\textipa{/i/} (written in Cyrillic as \foreignlanguage{russian}{и}), 
+such that the vowel phoneme \textipa{/i/} occurs after soft (i.e. 
+palatalized) consonants, whereas the vowel phoneme \textipa{/\textbari/} 
+occurs after hard (i.e. plain) consonants. Other linguists believe in a
+six-vowel analysis, asserting \textipa{/\textbari/} as a separate phoneme.
 
-\subsection{The Ruler}
-The ACL 2019 style defines a printed ruler which should be presented in the
-version submitted for review.  The ruler is provided in order that
-reviewers may comment on particular lines in the paper without
-circumlocution.  If you are preparing a document without the provided
-style files, please arrange for an equivalent ruler to
-appear on the final output pages.  The presence or absence of the ruler
-should not change the appearance of any other content on the page.  The
-camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment the {\small\verb|\aclfinalcopy|} command in the document preamble.)
+Another set of phonemes with disputed status in Russian are palatalized
+velar consonants (i.e. \textipa{/k\super j/}, \textipa{/g\super j/}, 
+and \textipa{/x\super j/}) as well as the palatalized voiceless alveolar
+sibilant affricate \textipa{/\t{ts}\super j/}. In most cases, the velar consonants
+become soft when followed by front vowels (except in the case of a word
+boundary between the consonant and vowel), and the affricate is generally
+always hard. It is only in certain loanwords and foreign names that
+the consonants may become palatalized in other contexts.
 
-Reviewers: note that the ruler measurements do not align well with
-lines in the paper -- this turns out to be very difficult to do well
-when the paper contains many figures and equations, and, when done,
-looks ugly. In most cases one would expect that the approximate
-location will be adequate, although you can also use fractional
-references (\emph{e.g.}, the first paragraph on this page ends at mark $108.5$).
+For the sake of our experiment, however, we have chosen to follow the
+five-vowel analysis and not count the extra palatalized consonants
+as Russian phonemes. The phonemic status of the vowel \textipa{/\textbari/}
+is marginal, only occuring isolated in the verb \foreignlanguage{russian}{ыкать}
+(to pronounce the sound \foreignlanguage{russian}{ы}) and in borrowed names
+of places in Russia. By following the five-vowel analysis, we can account for
+one less vowel, because \textipa{/\textbari/} and \textipa{/i/} are otherwise
+in complementary distribution in all other environments. Likewise, because
+the velar consonants are normally only palatalized following front vowels,
+there is no need to account for such redundant information.
 
-\subsection{Electronically-available resources}
-
-ACL provides this description in \LaTeX2e{} (\texttt{\small acl2019.tex}) and PDF
-format (\texttt{\small acl2019.pdf}), along with the \LaTeX2e{} style file used to
-format it (\texttt{\small acl2019.sty}) and an ACL bibliography style (\texttt{\small acl\_natbib.bst})
-and example bibliography (\texttt{\small acl2019.bib}).
-These files are all available at
-\texttt{\small http://acl2019.org/downloads/ acl2019-latex.zip}. 
- We
-strongly recommend the use of these style files, which have been
-appropriately tailored for the ACL 2019 proceedings.
-
-\subsection{Format of Electronic Manuscript}
-\label{sect:pdf}
-
-For the production of the electronic manuscript you must use Adobe's
-Portable Document Format (PDF). PDF files are usually produced from
-\LaTeX\ using the \textit{pdflatex} command. If your version of
-\LaTeX\ produces Postscript files, you can convert these into PDF
-using \textit{ps2pdf} or \textit{dvipdf}. On Windows, you can also use
-Adobe Distiller to generate PDF.
-
-Please make sure that your PDF file includes all the necessary fonts
-(especially tree diagrams, symbols, and fonts with Asian
-characters). When you print or create the PDF file, there is usually
-an option in your printer setup to include none, all or just
-non-standard fonts.  Please make sure that you select the option of
-including ALL the fonts. \textbf{Before sending it, test your PDF by
-  printing it from a computer different from the one where it was
-  created.} Moreover, some word processors may generate very large PDF
-files, where each page is rendered as an image. Such images may
-reproduce poorly. In this case, try alternative ways to obtain the
-PDF. One way on some systems is to install a driver for a postscript
-printer, send your document to the printer specifying ``Output to a
-file'', then convert the file to PDF.
-
-It is of utmost importance to specify the \textbf{A4 format} (21 cm
-x 29.7 cm) when formatting the paper. When working with
-\texttt{dvips}, for instance, one should specify \texttt{-t a4}.
-Or using the command \verb|\special{papersize=210mm,297mm}| in the latex
-preamble (directly below the \verb|\usepackage| commands). Then using 
-\texttt{dvipdf} and/or \texttt{pdflatex} which would make it easier for some.
-
-Print-outs of the PDF file on A4 paper should be identical to the
-hardcopy version. If you cannot meet the above requirements about the
-production of your electronic submission, please contact the
-publication chairs as soon as possible.
-
-\subsection{Layout}
-\label{ssec:layout}
-
-Format manuscripts two columns to a page, in the manner these
-instructions are formatted. The exact dimensions for a page on A4
-paper are:
-
-\begin{itemize}
-\item Left and right margins: 2.5 cm
-\item Top margin: 2.5 cm
-\item Bottom margin: 2.5 cm
-\item Column width: 7.7 cm
-\item Column height: 24.7 cm
-\item Gap between columns: 0.6 cm
-\end{itemize}
-
-\noindent Papers should not be submitted on any other paper size.
- If you cannot meet the above requirements about the production of 
- your electronic submission, please contact the publication chairs 
- above as soon as possible.
-
-\subsection{Fonts}
-
-For reasons of uniformity, Adobe's \textbf{Times Roman} font should be
-used. In \LaTeX2e{} this is accomplished by putting
-
-\begin{quote}
-\begin{verbatim}
-\usepackage{times}
-\usepackage{latexsym}
-\end{verbatim}
-\end{quote}
-in the preamble. If Times Roman is unavailable, use \textbf{Computer
-  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
-  10\% less dense than Adobe's Times Roman font.
-
-\begin{table}[t!]
-\begin{center}
-\begin{tabular}{|l|rl|}
-\hline \textbf{Type of Text} & \textbf{Font Size} & \textbf{Style} \\ \hline
-paper title & 15 pt & bold \\
-author names & 12 pt & bold \\
-author affiliation & 12 pt & \\
-the word ``Abstract'' & 12 pt & bold \\
-section titles & 12 pt & bold \\
-subsection titles & 11 pt & bold \\
-document text & 11 pt  &\\
-captions & 10 pt & \\
-abstract text & 10 pt & \\
-bibliography & 10 pt & \\
-footnotes & 9 pt & \\
-\hline
-\end{tabular}
-\end{center}
-\caption{\label{font-table} Font guide. }
-\end{table}
-
-\subsection{The First Page}
-\label{ssec:first}
-
-Center the title, author's name(s) and affiliation(s) across both
-columns. Do not use footnotes for affiliations. Do not include the
-paper ID number assigned during the submission process. Use the
-two-column format only when you begin the abstract.
-
-\textbf{Title}: Place the title centered at the top of the first page, in
-a 15-point bold font. (For a complete guide to font sizes and styles,
-see Table~\ref{font-table}) Long titles should be typed on two lines
-without a blank line intervening. Approximately, put the title at 2.5
-cm from the top of the page, followed by a blank line, then the
-author's names(s), and the affiliation on the following line. Do not
-use only initials for given names (middle initials are allowed). Do
-not format surnames in all capitals (\emph{e.g.}, use ``Mitchell'' not
-``MITCHELL'').  Do not format title and section headings in all
-capitals as well except for proper names (such as ``BLEU'') that are
-conventionally in all capitals.  The affiliation should contain the
-author's complete address, and if possible, an electronic mail
-address. Start the body of the first page 7.5 cm from the top of the
-page.
-
-The title, author names and addresses should be completely identical
-to those entered to the electronical paper submission website in order
-to maintain the consistency of author information among all
-publications of the conference. If they are different, the publication
-chairs may resolve the difference without consulting with you; so it
-is in your own interest to double-check that the information is
-consistent.
-
-\textbf{Abstract}: Type the abstract at the beginning of the first
-column. The width of the abstract text should be smaller than the
-width of the columns for the text in the body of the paper by about
-0.6 cm on each side. Center the word \textbf{Abstract} in a 12 point bold
-font above the body of the abstract. The abstract should be a concise
-summary of the general thesis and conclusions of the paper. It should
-be no longer than 200 words. The abstract text should be in 10 point font.
-
-\textbf{Text}: Begin typing the main body of the text immediately after
-the abstract, observing the two-column format as shown in the present document. Do not include page numbers.
-
-\textbf{Indent}: Indent when starting a new paragraph, about 0.4 cm. Use 11 points for text and subsection headings, 12 points for section headings and 15 points for the title.
-
-\begin{table}
-\centering
-\small
-\begin{tabular}{cc}
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Command} & \textbf{Output}\\\hline
-\verb|{\"a}| & {\"a} \\
-\verb|{\^e}| & {\^e} \\
-\verb|{\`i}| & {\`i} \\ 
-\verb|{\.I}| & {\.I} \\ 
-\verb|{\o}| & {\o} \\
-\verb|{\'u}| & {\'u}  \\ 
-\verb|{\aa}| & {\aa}  \\\hline
-\end{tabular} & 
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Command} & \textbf{Output}\\\hline
-\verb|{\c c}| & {\c c} \\ 
-\verb|{\u g}| & {\u g} \\ 
-\verb|{\l}| & {\l} \\ 
-\verb|{\~n}| & {\~n} \\ 
-\verb|{\H o}| & {\H o} \\ 
-\verb|{\v r}| & {\v r} \\ 
-\verb|{\ss}| & {\ss} \\\hline
-\end{tabular}
-\end{tabular}
-\caption{Example commands for accented characters, to be used in, \emph{e.g.}, \BibTeX\ names.}\label{tab:accents}
-\end{table}
-
-\subsection{Sections}
-
-\textbf{Headings}: Type and label section and subsection headings in the
-style shown on the present document.  Use numbered sections (Arabic
-numerals) in order to facilitate cross references. Number subsections
-with the section number and the subsection number separated by a dot,
-in Arabic numerals.
-Do not number subsubsections.
+Table 2 shows the scheme that we used to categorize Russian phonemes
+on five simple dimensions.
 
@@
 \centering
-\begin{tabular}{lll}
-  output & natbib & previous ACL style files\\
-  \hline
-  \citep{Gusfield:97} & \verb|\citep| & \verb|\cite| \\
-  \citet{Gusfield:97} & \verb|\citet| & \verb|\newcite| \\
-  \citeyearpar{Gusfield:97} & \verb|\citeyearpar| & \verb|\shortcite| \\
+\begin{tabular}{cccccccccccccc}
+    \multicolumn{2}{c}{} & \multicolumn{12}{c}{Place}\\\hline
+    \multicolumn{2}{c}{} & \multicolumn{4}{c}{Front} & \multicolumn{4}{c}{Middle} & \multicolumn{4}{c}{Back}\\\hline
+    \multicolumn{2}{c}{} & \multicolumn{2}{c}{Voiceless} & \multicolumn{2}{c}{Voiced} & \multicolumn{2}{c}{Voiceless} & \multicolumn{2}{c}{Voiced} & \multicolumn{2}{c}{Voiceless} & \multicolumn{2}{c}{Voiced}\\\hline
+    \multicolumn{2}{c}{} & H & S & H & S & H & S & H & S & H & S & H & S\\\hline
+    \multirow{2}{50px}{Interrupted} & Stop & p & p' & b & b' & t & t' & d & d' & k & - & g & -\\
+    & Nasal & - & - & m & m' & - & - & n & n' & - & - & - & - \\
+    \multirow{2}{50px}{Cont.Cons.} & Fricative & f & f' & v & v' & s & s' & z & z' & s\` & t\_s\textbackslash & z\` & -\\
+                                        & Liq/Sem. & - & - & l & l' & t\_s & - & r & r' & x & - & - & j\\
+    \multirow{2}{50px}{Vowel} & Close & - & - & i & - & - & - & - & - & - & - & u & -\\
+                              & Open & - & - & e & - & - & - & a & - & - & - & o & -\\
 \end{tabular}
-\caption{Citation commands supported by the style file.
-  The citation style is based on the natbib package and
-  supports all natbib citation commands.
-  It also supports commands defined in previous ACL style files
-  for compatibility.
-  }
+\caption{Categorization of Russian phonemes on five dimensions.
+All phonemes are written using X-SAMPA.}
 \end{table*}
+
+\section{Experimental setup}
+
+To run our experiment, we first obtained a Russian word list, filtered the
+word data, then preprocessed the data into a format that can be used by a
+neural network model. We then trained our neural network models based on
+our word data and evaluated the results. The procedure for these steps
+is outlined in the subsections below.
+
+\subsection{Collecting Russian nouns and declensions}
+
+In order to train the neural network, we needed a list of Russian nouns
+and their genitive singular and plural forms. For this, we scraped words
+from Wiktionary. Wiktionary is an online, multilingual, and free project
+to create a dictionary for all natural languages. These dictionary entries
+often contain additional information about words in languages, such as verb
+conjugations, example sentences, synonyms and antonyms, and most importantly,
+noun declensions.
+
+Every Russian noun with an entry on Wiktionary is accessible through the category
+page \textbf{Category:Russian Nouns}. Using Wiktionary's MediaWiki API, we gathered
+every noun in this category in batches of 500 (due to API limitations). For each
+batch, we checked to make sure that the word only contained letters in the modern
+Russian Cyrillic alphabet, because Wiktionary also has entries for Russian words
+containing characters from different alphabets (e.g. \textbf{\foreignlanguage{russian}{QR-код}},
+meaning ``QR code'') as well as alternate spellings of words containing archaic letters 
+unused since the Russian spelling reform of 1918.
+
+After performing this initial pruning, we then queried the API for each word's
+entry in Wiktionary and attempted to extract its genitive singular and plural
+forms from the declension table available in the entry. If either the singular
+or plural genitive forms were missing from the table, the noun is discarded.
+This can happen if a word is an indeclinable loanword or only exists in either
+the singular or plural forms. We then save the word, its genitive singular and
+plural forms into a wordlist, which will be consulted during the pre-processing
+and training phases.
+
+\subsection{Noun data pre-processing}
+
+Before the noun data can be used as input to train the neural network, it must
+first be converted into a representation usable by the model.
+
+The first transformation that takes place is converting the noun forms'
+Russian Cyrillic representation into a phonemic representation in X-SAMPA\@.
+Because the Russian alphabet is much more phonemically consistent than English,
+meaning it is possible to derive a word's phonemic representation by iterating
+through its Cyrillic representation one letter at a time.
+
+% TODO: explain this better
+After being converted into X-SAMPA, each phoneme would then be converted into
+its Wickelphone/Wickelfeature representations.
+
+The Wickelfeatures would then be converted into arrays of activations for
+certain features.
+
+\subsection{Neural network}
+
+The basis of our neural network is a gradient-descent neural network powered by the
+\emph{PyTorch} library. It uses a single linear transformation layer with the
+input size being the number of Wickelfeatures that can be activated for,
+and the output size being either 1 or 2 times the number of Wickelfeatures
+that can be activated for, depending on whether the model is one of the
+two smaller models learning individual genitive cases or the larger model
+learning both genitive cases.
 
@@
 as sentence constituents.
-
-We suggest that instead of
-\begin{quote}
-  ``\cite{Gusfield:97} showed that ...''
-\end{quote}
-you use
-\begin{quote}
-``Gusfield \shortcite{Gusfield:97}   showed that ...''
-\end{quote}
-
-If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
-can use the command \verb|\citet| (cite in text)
-to get ``author (year)'' citations.
-
-You can use the command \verb|\citealp| (alternative cite without 
-parentheses) to get ``author year'' citations (which is useful for 
-using citations within parentheses, as in \citealp{Gusfield:97}).
-
-If the Bib\TeX{} file contains DOI fields, the paper
-title in the references section will appear as a hyperlink
-to the DOI, using the hyperref \LaTeX{} package.
-To disable the hyperref package, load the style file
-with the \verb|nohyperref| option: \\{\small
-\verb|\usepackage[nohyperref]{acl2019}|}
-
-\textbf{Compilation Issues}: Some of you might encounter the following error during compilation:
-
-``{\em \verb|\pdfendlink| ended up in different nesting level than \verb|\pdfstartlink|.}''
-
-This happens when \verb|pdflatex| is used and a citation splits across a page boundary.
-To fix this, the style file contains a patch consisting of the following two lines: (1) \verb|\RequirePackage{etoolbox}| (line 454 in \texttt{acl2019.sty}), and (2) A long line below (line 455 in \texttt{acl2019.sty}).
-
-If you still encounter compilation issues even with the patch enabled, disable the patch by commenting the two lines, and then disable the \verb|hyperref| package (see above), recompile and see the problematic citation.
-Next rewrite that sentence containing the citation. (See, {\em e.g.}, {\small\tt http://tug.org/errors.html})
-
-\textbf{Digital Object Identifiers}:  As part of our work to make ACL
-materials more widely used and cited outside of our discipline, ACL
-has registered as a CrossRef member, as a registrant of Digital Object
-Identifiers (DOIs), the standard for registering permanent URNs for
-referencing scholarly materials.  As of 2017, we are requiring all
-camera-ready references to contain the appropriate DOIs (or as a
-second resort, the hyperlinked ACL Anthology Identifier) to all cited
-works.  Thus, please ensure that you use Bib\TeX\ records that contain
-DOI or URLs for any of the ACL materials that you reference.
-Appropriate records should be found for most materials in the current
-ACL Anthology at \url{http://aclanthology.info/}.
-
-As examples, we cite \cite{P16-1001} to show you how papers with a DOI
-will appear in the bibliography.  We cite \cite{C14-1001} to show how
-papers without a DOI but with an ACL Anthology Identifier will appear
-in the bibliography.
-
-As reviewing will be double-blind, the submitted version of the papers
-should not include the authors' names and affiliations. Furthermore,
-self-references that reveal the author's identity, \emph{e.g.},
-\begin{quote}
-``We previously showed \cite{Gusfield:97} ...''  
-\end{quote}
-should be avoided. Instead, use citations such as 
-\begin{quote}
-``\citeauthor{Gusfield:97} \shortcite{Gusfield:97}
-previously showed ... ''
-\end{quote}
-
-Any preliminary non-archival versions of submitted papers should be listed in the submission form but not in the review version of the paper.
-ACL 2019 reviewers are generally aware that authors may present preliminary versions of their work in other venues, but will not be provided the list of previous presentations from the submission form.
-
-\textbf{Please do not use anonymous citations} and do not include
- when submitting your papers. Papers that do not
-conform to these requirements may be rejected without review.
-
-\textbf{References}: Gather the full set of references together under
-the heading \textbf{References}; place the section before any Appendices. 
-Arrange the references alphabetically
-by first author, rather than by order of occurrence in the text.
-By using a .bib file, as in this template, this will be automatically 
-handled for you. See the \verb|\bibliography| commands near the end for more.
-
-Provide as complete a citation as possible, using a consistent format,
-such as the one for \emph{Computational Linguistics\/} or the one in the 
-\emph{Publication Manual of the American 
-Psychological Association\/}~\cite{APA:83}. Use of full names for
-authors rather than initials is preferred. A list of abbreviations
-for common computer science journals can be found in the ACM 
-\emph{Computing Reviews\/}~\cite{ACM:83}.
-
-The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
-American Psychological Association format, allowing regular citations, 
-short citations and multiple citations as described above.
-
-\begin{itemize}
-\item Example citing an arxiv paper: \cite{rasooli-tetrault-2015}. 
-\item Example article in journal citation: \cite{Ando2005}.
-\item Example article in proceedings, with location: \cite{borsch2011}.
-\item Example article in proceedings, without location: \cite{andrew2007scalable}.
-\end{itemize}
-See corresponding .bib file for further details.
-
-Submissions should accurately reference prior and related work, including code and data.
-If a piece of prior work appeared in multiple venues, the version that appeared in a refereed, archival venue should be referenced.
-If multiple versions of a piece of prior work exist, the one used by the authors should be referenced.
-Authors should not rely on automated citation indices to provide accurate references for prior and related work.
-
-\textbf{Appendices}: Appendices, if any, directly follow the text and the
-references (but see above).  Letter them in sequence and provide an
-informative title: \textbf{Appendix A. Title of Appendix}.
-
-\subsection{Footnotes}
-
-\textbf{Footnotes}: Put footnotes at the bottom of the page and use 9
-point font. They may be numbered or referred to by asterisks or other
-symbols.\footnote{This is how a footnote should appear.} Footnotes
-should be separated from the text by a line.\footnote{Note the line
-separating the footnotes from the text.}
-
-\subsection{Graphics}
-
-\textbf{Illustrations}: Place figures, tables, and photographs in the
-paper near where they are first discussed, rather than at the end, if
-possible.  Wide illustrations may run across both columns.  Color
-illustrations are discouraged, unless you have verified that  
-they will be understandable when printed in black ink.
-
-\textbf{Captions}: Provide a caption for every illustration; number each one
-sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
-Caption of the Table.''  Type the captions of the figures and 
-tables below the body, using 10 point text.
-Captions should be placed below illustrations.
-Captions that are one line are centered (see Table~\ref{font-table}).
-Captions longer than one line are left-aligned (see Table~\ref{tab:accents}).
-Do not overwrite the default caption sizes.
-The acl2019.sty file is compatible with the caption and subcaption packages; do not add optional arguments.
-
-\subsection{Accessibility}
-\label{ssec:accessibility}
-
-In an effort to accommodate people who are color-blind (as well as those printing
-to paper), grayscale readability for all accepted papers will be
-encouraged.  Color is not forbidden, but authors should ensure that
-tables and figures do not rely solely on color to convey critical
-distinctions. A simple criterion: All curves and points in your figures should be clearly distinguishable without color.
-
-% Min: no longer used as of ACL 2018, following ACL exec's decision to
-% remove this extra workflow that was not executed much.
-% BEGIN: remove
-%% \section{XML conversion and supported \LaTeX\ packages}
-
-%% Following ACL 2014 we will also we will attempt to automatically convert 
-%% your \LaTeX\ source files to publish papers in machine-readable 
-%% XML with semantic markup in the ACL Anthology, in addition to the 
-%% traditional PDF format.  This will allow us to create, over the next 
-%% few years, a growing corpus of scientific text for our own future research, 
-%% and picks up on recent initiatives on converting ACL papers from earlier 
-%% years to XML.
-
-%% We encourage you to submit a ZIP file of your \LaTeX\ sources along
-%% with the camera-ready version of your paper. We will then convert them
-%% to XML automatically, using the LaTeXML tool
-%% (\url{http://dlmf.nist.gov/LaTeXML}). LaTeXML has \emph{bindings} for
-%% a number of \LaTeX\ packages, including the ACL 2018 stylefile. These
-%% bindings allow LaTeXML to render the commands from these packages
-%% correctly in XML. For best results, we encourage you to use the
-%% packages that are officially supported by LaTeXML, listed at
-%% \url{http://dlmf.nist.gov/LaTeXML/manual/included.bindings}
-% END: remove
-
-\section{Translation of non-English Terms}
-
-It is also advised to supplement non-English characters and terms
-with appropriate transliterations and/or translations
-since not all readers understand all such characters and terms.
-Inline transliteration or translation can be represented in
-the order of: original-form transliteration ``translation''.
-
-\section{Length of Submission}
-\label{sec:length}
-
-The ACL 2019 main conference accepts submissions of long papers and
-short papers.
- Long papers may consist of up to eight (8) pages of
-content plus unlimited pages for references. Upon acceptance, final
-versions of long papers will be given one additional page -- up to nine (9)
-pages of content plus unlimited pages for references -- so that reviewers' comments
-can be taken into account. Short papers may consist of up to four (4)
-pages of content, plus unlimited pages for references. Upon
-acceptance, short papers will be given five (5) pages in the
-proceedings and unlimited pages for references. 
-For both long and short papers, all illustrations and tables that are part
-of the main text must be accommodated within these page limits, observing
-the formatting instructions given in the present document. Papers that do not conform to the specified length and formatting requirements are subject to be rejected without review.
-
-ACL 2019 does encourage the submission of additional material that is relevant to the reviewers but not an integral part of the paper.
-There are two such types of material: appendices, which can be read, and non-readable supplementary materials, often data or code.
-Do not include this additional material in the same document as your main paper.
-Additional material must be submitted as one or more separate files, and must adhere to the same anonymity guidelines as the main paper.
-The paper must be self-contained: it is optional for reviewers to look at the supplementary material.
-Papers should not refer, for further detail, to documents, code or data resources that are not available to the reviewers.
-Refer to Appendix~\ref{sec:appendix} and Appendix~\ref{sec:supplemental} for further information.
-
-Workshop chairs may have different rules for allowed length and
-whether supplemental material is welcome. As always, the respective
-call for papers is the authoritative source.
 
@@
 
-\section{Supplemental Material}
-\label{sec:supplemental}
-Submissions may include non-readable supplementary material used in the work and described in the paper.
-Any accompanying software and/or data should include licenses and documentation of research review as appropriate.
-Supplementary material may report preprocessing decisions, model parameters, and other details necessary for the replication of the experiments reported in the paper.
-Seemingly small preprocessing decisions can sometimes make a large difference in performance, so it is crucial to record such decisions to precisely characterize state-of-the-art methods.
-
-Nonetheless, supplementary material should be supplementary (rather
-than central) to the paper. \textbf{Submissions that misuse the supplementary 
-material may be rejected without review.}
-Supplementary material may include explanations or details
-of proofs or derivations that do not fit into the paper, lists of
-features or feature templates, sample inputs and outputs for a system,
-pseudo-code or source code, and data. (Source code and data should
-be separate uploads, rather than part of the paper).
-
-The paper should not rely on the supplementary material: while the paper
-may refer to and cite the supplementary material and the supplementary material will be available to the
-reviewers, they will not be asked to review the
-supplementary material.
-
 \end{document}

### 2023-12-10 paper/paper.tex
update tex
panic model adjustments, paper additions

@@
 \newcommand\BibTeX{B\textsc{ib}\TeX}
+
+\newcommand{\citenp}[1]{\citeauthor{#1} (\citeyear{#1})}
 
@@
 language acquisition devices (LAD) in order to model how children acquire
-the English past tense. Most notably, Rumelhart and McClelland (1986)
+the English past tense. Most notably, \citenp{RMcC}
 made use of a perceptron model to model the learning of English past
@@
 by changing the ending of the word, and the exact manner by which an ending
-changes depends on the ending, case, and number of the noun. Table 1 shows 
+changes depends on the ending, case, and number of the noun. Table \ref{declension} shows 
 the different possible endings of the Russian nominative and genitive cases in 
@@
 
-Rumelhart and McClelland (1986) made use of a scheme to encode
+\citenp{RMcC} made use of a scheme to encode
 English phoneme units, known as Wickelphones and Wickelfeatures, described
-in Wickelgren (1969). Wickelphones are phoneme units representing
+in \citenp{Wickelgren}. Wickelphones are phoneme units representing
 each phone in a word as triples, which consist of the phoneme itself,
@@
 
-Table 2 shows the scheme that we used to categorize Russian phonemes
+Table \ref{phonemes} shows the scheme that we used to categorize Russian phonemes
 on five simple dimensions.
@@
 \end{tabular}
-\caption{Categorization of Russian phonemes on five dimensions.
+\caption{\label{phonemes} Categorization of Russian phonemes on five dimensions.
 All phonemes are written using X-SAMPA.}
@@
 
-% TODO: explain this better
-After being converted into X-SAMPA, each phoneme would then be converted into
-its Wickelphone/Wickelfeature representations.
+However, phonemic representations are not specific enough for the neural network.
+As previously referenced, \citenp{RMcC} refer to \citenp{Wickelgren}
+to devise a featurization system using Wickelphones and Wickelfeatures.
+In particular, each Wickelphone has a set of at most 25 Wickelfeatures that its presence activates,
+which are derived from the phonological features of all three phonemes present in the Wickelphone.
 
-The Wickelfeatures would then be converted into arrays of activations for
-certain features.
+For all phonemes other than the special word-boundary marker (which has a single unique `boundary' feature),
+each phoneme has five features that correspond to the dimensions outlined in Table \ref{phonemes}.
+For each feature in the central phoneme of the Wickelphone, at most five Wickelfeatures are generated:
+each feature in the left context phoneme, the selected feature of the central phoneme,
+and each feature in the right context. An example of this for the Wickelphone
+\textbf{\textsubscript{j}a\textsubscript{z}} is present in Table \ref{features}.
+Computationally, this algorithm corresponds well to a series of vector repetitions and a matrix transposition.
+
+Each of these activated Wickelfeatures are then ``blurred": for a Wickelfeature $\left< f_1, f_2, f_3 \right>$,
+all Wickelfeatures of the form $\left< * , f_2, f_3 \right>$ and $\left< f_1, f_2, * \right>$ have a random chance of being activated
+as well. \citenp{RMcC} claim that this improves the neural network's ability to generalize between
+phonological features.
+
+However, to avoid computational blowup, Wickelfeatures with different dimensions for their left and right features
+(other than the special word boundary feature) are filtered out of the computation. This brings the total number
+of possible Wickelfeatures down from 2028 to only 660.
+
+Activations across Wickelfeatures are thus represented by a boolean 660-vector, where a 1 in index $k$ corresponds to
+Wickelfeature $k$ being activated. A word's total activation is all the Wickelfeatures activated by the word's composite Wickelphones. This is easily implemented with a vector sum operation.
+
+\begin{table}[t!]
+\begin{center}
+	\small
+\begin{tabular}{r|c|c|c}
+	\# & Left Context & Central Phoneme & Right Context \\ \hline
+	1 & Continuous & Vowel & Continuous \\
+	2 & Liquid & Vowel & Fricative \\
+	3 & Back & Vowel & Middle \\
+	4 & Voiced & Vowel & Voiced \\
+	5 & Soft & Vowel & Hard \\
+	6 & Continuous & Open & Continuous \\
+	7 & Liquid & Open & Fricative \\
+	8 & Back & Open & Middle \\
+	9 & Voiced & Open & Voiced \\
+	10 & Soft & Open & Hard \\
+	11 & Continuous & Middle & Continuous \\
+	12 & Liquid & Middle & Fricative \\
+	13 & Back & Middle & Middle \\
+	14 & Voiced & Middle & Voiced \\
+	15 & Soft & Middle & Hard \\
+	16 & Continuous & Voiced & Continuous \\
+	17 & Liquid & Voiced & Fricative \\
+	18 & Back & Voiced & Middle \\
+	19 & Voiced & Voiced & Voiced \\
+	20 & Soft & Voiced & Hard \\
+	21 & Continuous & Hard & Continuous \\
+	22 & Liquid & Hard & Fricative \\
+	23 & Back & Hard & Middle \\
+	24 & Voiced & Hard & Voiced \\
+	25 & Soft & Hard & Hard \\
+\end{tabular}
+\end{center}
+\caption{\label{features} The Wickelfeatures activated by the Wickelphone
+		                  \textbf{\textsubscript{j}a\textsubscript{z}}.}
+\end{table}
 
@@
 
-\textbf{Citations}: Citations within the text appear in parentheses
-as~\cite{Gusfield:97} or, if the author's name appears in the text
-itself, as Gusfield~\shortcite{Gusfield:97}.
-Using the provided \LaTeX\ style, the former is accomplished using
-{\small\verb|\cite|} and the latter with {\small\verb|\shortcite|} or {\small\verb|\newcite|}.
-Collapse multiple citations as in~\cite{Gusfield:97,Aho:72}; this is accomplished with the provided style using commas within the {\small\verb|\cite|} command, \emph{e.g.}, {\small\verb|\cite{Gusfield:97,Aho:72}|}.
-Append lowercase letters to the year in cases of ambiguities.
- Treat double authors as
-in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two
-authors are involved. Collapse multiple citations as
-in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations
-as sentence constituents.
+During training, the network's input is the activation vector for the nominative singular form
+of a noun, and the expected output is the activation vector for that noun's genitive form
+(singular or plural, depending on the network being trained). However, the larger model attempting
+to learn both cases at once simply has the singular and plural genitive activation vectors concatenated together,
+with the intent being that the model will eventually output two ``words" at once.
 
-\section*{Acknowledgments}
+\subsection{Decoding model output}
+Described in the Appendix of their paper, \citenp{RMcC} use what they term a ``binding network" to decode the feature activation vectors
+outputted by the learning network into a full phonological representation. Because the activation
+vector encoding scheme has no representation of the temporal dimension of a word (i.e., it does
+not directly encode \textit{where} features and source phonemes occur in the source word), a separate
+decoder is necessary. The binding network does not learn its weights, but rather each weight is determined
+by the strength of its previous output ``in time", proportional to the probability of the corresponding feature
+being activated (i.e., the network output).
 
-The acknowledgments should go immediately before the references.  Do
-not number the acknowledgments section. Do not include this section
-when submitting your paper for review. \\
+The network attempts to model how each Wickelphone contributes to the Wickelfeatures present in the activation
+vector, with the idea being ``to find a set of output features [Wickelphones] that accounts for as many as possible of the output features while minimizing the number of input features [Wickelfeatures] accounted for by more than one output feature.
+Thus, we want each of the output features to \textit{compete} for input features" \cite[pp.
+269]{RMcC}.
 
-\noindent \textbf{Preparing References:} \\
-Include your own bib file like this:
-\verb|\bibliographystyle{acl_natbib}|
-\verb|\bibliography{acl2019}|
+\section{Results}
+Unfortunately, our results were decidedly inconclusive.
+The neural networks did successfully learn during training,
+with a final loss of approximately 0.1 (down from 0.16 after 100 epochs).
+However, the decoder network did not work as hoped, and only output two Wickelphones
+when given output from any of the neural networks. For example, when given the network predicted activations
+for the input \foreignlanguage{russian}{язык} (X-SAMPA \textbf{jazik}), the decoder only ever output
+the Wickelphones \textbf{\textsubscript{\#}j\textsubscript{a}} and \textbf{\textsubscript{j}a\textsubscript{\#}};
+for the input \foreignlanguage{russian}{жена} (X-SAMPA \textbf{z'ena}), the decoder only returned \textbf{\textsubscript{\#}z'\textsubscript{e}} and \textbf{\textsubscript{z'}e\textsubscript{\#}}.
 
-where \verb|acl2019| corresponds to a acl2019.bib file.
-\bibliography{acl2019}
+\subsection{Discussion}
+These results seem to suggest that a featural representation of words that properly preserves temporal
+information is necessary to get meaningful output from these models. It is also possible that the
+specification for the binding network in \citenp{RMcC} was overly general, and thus some nuance in implementation
+may have been missed.
+
+\section{Conclusion}
+
+\bibliography{paper}{}
 \bibliographystyle{acl_natbib}

### 2023-12-10 binding.py

@@ -17,7 +17,7 @@ for _, nom, gen_sg, gen_pl in build_word_data():
         all_phones.add(phone)
     for phone in gen_pl:
         all_phones.add(phone)
-all_phones = list(all_phones)
+all_phones = sorted(list(all_phones)) # sort to ensure deterministic ordering, since sets are unordered
 
 input_features = len(wickelfeatures)
 output_features = len(all_phones)
@@ -39,11 +39,29 @@ WORD_SIZE = 10
 def predict_word(prediction):
     i = np.array(prediction)
 
+    import time
+
+    output_phones: list[WickelPhone] = []
+
     output = np.zeros([WORD_SIZE, output_features])
     for t in range(WORD_SIZE):
+        start = time.time()
+
         prev_time = np.ones(output_features) if t < 1 else output[t - 1]
 
         for k in range(output_features):
+            this_phone = all_phones[k]
+            this_left, this_center, _ = this_phone.phones
+            if not output_phones:
+                if this_left != "#":
+                    output[t, k] = 0
+                    continue
+            else:
+                _, prev_center, prev_right = output_phones[-1].phones
+                if this_left != prev_center or this_center != prev_right:
+                    output[t, k] = 0
+                    continue
+
             js = filtered_activations[k]
             ls = feature_to_phones[js]
             ls = functools.reduce(lambda s, t: s | t, ls)
@@ -53,9 +71,20 @@ def predict_word(prediction):
             denominator = np.sum(prev_time[ls])
 
             output[t, k] = numerator / denominator
-        print(f"calculated all of time {t}")
 
-    return output
+            print(f"{k + 1}/{output_features}", end="\r")
+        
+        end = time.time()
+
+        next_phone = all_phones[np.argmax(output[t])]
+
+        output_phones.append(next_phone)
+        # print(f"predicted phone {t + 1} in {end - start:.2f}s - {next_phone}")
+
+        if next_phone.phones[2] == '#':
+            break
+
+    return output_phones
 
 
 if __name__ == "__main__":
@@ -67,8 +96,6 @@ if __name__ == "__main__":
     with open("decoded.json", "w") as f:
         out = {}
         for k, v in p.items():
-            decoded = predict_word(v)
-            out[k] = [str(all_phones[i]) for i in np.argmax(decoded, axis=1)]
+            out[k] = [str(p) for p in predict_word(v)]
             print(f"{k}: {out[k]}")
-            out[k + "_weights"] = decoded.tolist()
         json.dump(out, f)

### 2023-12-10 model.py

@@ -7,7 +7,13 @@ import torch.nn as nn
 import torch.optim as optim
 
 from xsampa import cyrillic_to_xsampa
-from wickel import WickelPhone, xsampa_to_phones, wickelfeatures, feature_index_map, build_word_data
+from wickel import (
+    WickelPhone,
+    xsampa_to_phones,
+    wickelfeatures,
+    feature_index_map,
+    build_word_data,
+)
 
 
 def activations_for_phone(phone: WickelPhone):
@@ -49,15 +55,21 @@ class Net(nn.Module):
         return x
 
 
-net_gen_sg = Net()
-net_gen_pl = Net()
-net_both = Net(output_multiplier=2)
+if torch.cuda.is_available():
+    dev = "cuda"
+else:
+    dev = "cpu"
+device = torch.device(dev)
 
-loss_criterion = nn.L1Loss()
+net_gen_sg = Net().to(device)
+net_gen_pl = Net().to(device)
+net_both = Net(output_multiplier=2).to(device)
+
+loss_criterion = nn.L1Loss().to(device)
 
 
 def run_training(model, mapping_pairs, num_epochs):
-    optimizer = optim.SGD(model.parameters(), lr=10.0)
+    optimizer = optim.SGD(model.parameters(), lr=15.0)
 
     for epoch in range(num_epochs):
         for input_rep, target_rep in mapping_pairs:
@@ -90,17 +102,26 @@ def build_activations():
             print(f"built activations for {word}")
 
             words[word] = {
-                "nom_sg": nom_act.tolist(),
-                "gen_sg": sg_act.tolist(),
-                "gen_pl": pl_act.tolist(),
+                "nom_sg": {
+                    "phones": [str(p) for p in nom],
+                    "activations": nom_act.tolist(),
+                },
+                "gen_sg": {
+                    "phones": [str(p) for p in sg],
+                    "activations": sg_act.tolist(),
+                },
+                "gen_pl": {
+                    "phones": [str(p) for p in pl],
+                    "activations": pl_act.tolist(),
+                },
             }
         with open("activations.json", "w") as f:
             json.dump(words, f)
 
     for acts in words.values():
-        yield torch.FloatTensor(acts["nom_sg"]), torch.FloatTensor(
-            acts["gen_sg"]
-        ), torch.FloatTensor(acts["gen_pl"])
[…]

### 2023-12-10 new_decoder.py

@@ -1,0 +1,125 @@
+import json
+import os.path
+
+import numpy as np
+import torch
+import torch.nn as nn
+import torch.optim as optim
+
+from wickel import (
+    WickelPhone,
+    build_word_data,
+    wickelfeatures,
+)
+
+all_phones = set[WickelPhone]()
+for _, nom, gen_sg, gen_pl in build_word_data():
+    for phone in nom:
+        all_phones.add(phone)
+    for phone in gen_sg:
+        all_phones.add(phone)
+    for phone in gen_pl:
+        all_phones.add(phone)
+all_phones = sorted(
+    list(all_phones)
+)  # sort to ensure deterministic ordering, since sets are unordered
+phones_index_map = { str(p): i for i, p in enumerate(all_phones) }
+
+
+class Decoder(nn.Module):
+    def __init__(self):
+        super(Decoder, self).__init__()
+
+        self.layer = nn.Linear(
+            in_features=len(wickelfeatures) + len(all_phones),
+            out_features=len(all_phones),
+            bias=True,
+        )
+        self.activation = nn.Sigmoid()
+
+    def forward(self, x):
+        x = self.layer(x)
+        x = self.activation(x)
+        return x
+
+
+if torch.cuda.is_available():
+    dev = "cuda"
+else:
+    dev = "cpu"
+device = torch.device(dev)
+
+dec_model = Decoder().to(device)
+
+loss_criterion = nn.L1Loss().to(device)
+
+
+def run_training(model, mapping_pairs, num_epochs):
+    optimizer = optim.SGD(model.parameters(), lr=0.1)
+
+    for epoch in range(num_epochs):
+        for i, (input_rep, target_rep) in enumerate(mapping_pairs):
+            input_rep = input_rep.to(device).view(1, -1)
+            target_rep = target_rep.to(device).view(1, -1)
+
+            optimizer.zero_grad()
+            output = model(input_rep)
+            loss = loss_criterion(output, target_rep)
+            loss.backward()
+            optimizer.step()
+            print(f"trained {i + 1}/{len(mapping_pairs)}", end="\r")
+
+        print(f"loss at epoch {epoch}: {loss.item():.6f}")
+
+
+def build_training_data():
+    if not os.path.exists("activations.json"):
+        import model
+        model.build_activations()
+
+    with open("activations.json") as f:
+        words = json.load(f)
+    
+    for cases in words.values():
+        for case in cases.values():
+            last_phone = torch.zeros([len(all_phones)])
+
[…]

### 2023-12-10 wickel.py

@@ -1,4 +1,5 @@
 import csv
+import functools
 import itertools
 import random
 import os.path
@@ -161,9 +162,13 @@ wickelfeatures = list(all_wickelfeatures())
 feature_index_map = {f: i for i, f in enumerate(wickelfeatures)}
 
 
+@functools.total_ordering
 class WickelPhone:
     def __init__(self, first: str, second: str, third: str):
         self.phones = [first, second, third]
+
+    def __repr__(self):
+        return "{" + self.phones[0] + "}" + self.phones[1] + "{" + self.phones[2] + "}"
 
     def __hash__(self):
         return hash(self.__repr__())
@@ -173,10 +178,10 @@ class WickelPhone:
     
     def __str__(self):
         return self.__repr__()
-
-    def __repr__(self):
-        return "{" + self.phones[0] + "}" + self.phones[1] + "{" + self.phones[2] + "}"
-
+    
+    def __lt__(self, other):
+        return str(self) < str(other)
+    
     def activating_features(self):
         pre, central, post = [np.array(list(SegmentFeatures(p))) for p in self.phones]
 
@@ -215,9 +220,9 @@ def build_word_data():
         for row in csv.DictReader(file):
             yield (
                 row["title"],
-                xsampa_to_phones(cyrillic_to_xsampa(row["title"])),
-                xsampa_to_phones(cyrillic_to_xsampa(row["gen_sg"])),
-                xsampa_to_phones(cyrillic_to_xsampa(row["gen_pl"])),
+                list(xsampa_to_phones(cyrillic_to_xsampa(row["title"]))),
+                list(xsampa_to_phones(cyrillic_to_xsampa(row["gen_sg"]))),
+                list(xsampa_to_phones(cyrillic_to_xsampa(row["gen_pl"]))),
             )
 
 

### 2023-12-10 paper/paper.tex
add the github
conclusion added

@@
 
-%\aclfinalcopy % Uncomment this line for the final submission
+\aclfinalcopy % Uncomment this line for the final submission
 %\def\aclpaperid{***} %  Enter the acl Paper ID here
@@
 
-\title{Instructions for ACL 2019 Proceedings}
+\title{Modeling the Learning of the Russian Genitive Case}
 
@@
   J. L. McClelland.
+
+  All code written for this paper, including X-SAMPA translation and model training,
+  is available at \url{https://github.com/nickbclifford/genitive}.
 
@@
 
+Although we were unable to adequately test our neural network hypotheses due to the
+failure of our decoder network in converting the networks' outputs back to a phonological
+representation, we have nonetheless shown that the networks are capable of learning the
+genitive case's patterns to some extent with the Wickelphone/Wickelfeature schema described
+in this paper. Further investigation of the intended implementation of the decoder network
+could potentially demonstrate definitively whether or not the neural networks truly acquired 
+the proper formation of the genitive case, and which of the two sets of networks more closely
+models the acquisition of these cases in children with Russian as a native language.
+
+Future work in this subject matter could use our design to model learning with the other cases 
+of Russian, namely the dative, instrumental, and prepositional cases. It could also explore the
+acquisition of the different forms of adjectives, which also decline for grammatical case and
+number, or potentially the outcome of trying to learn both forms at once. All of these cases
+tend to show greater degrees of regularity compared to the genitive case, meaning that a neural network
+learning one or more of these cases is more likely to yield better results. This design
+may not suit modeling of learning the Russian accusative case, however, because the endings used in the accusative depends on whether the noun is animate (i.e. a living being) or not.
+
 \bibliography{paper}{}
 \bibliographystyle{acl_natbib}
-
-\appendix
-
-\section{Appendices}
-\label{sec:appendix}
-Appendices are material that can be read, and include lemmas, formulas, proofs, and tables that are not critical to the reading and understanding of the paper. 
-Appendices should be \textbf{uploaded as supplementary material} when submitting the paper for review. Upon acceptance, the appendices come after the references, as shown here. Use
-\verb|\appendix| before any appendix section to switch the section
-numbering over to letters.
-
 \end{document}

### 2023-12-10 new_decoder.py
fix decoder thing

@@ -95,11 +95,14 @@ def build_training_data():
                 last_phone = phone_pattern
 
 if __name__ == "__main__":
-    print("beginning training")
-    data = list(build_training_data())
-
-    run_training(dec_model, data, 15)
-    torch.save(dec_model.state_dict(), "decoder.pt")
+    if os.path.exists("decoder.pt"):
+        dec_model.load_state_dict(torch.load("decoder.pt"))
+        dec_model.eval()
+    else:
+        print("beginning training")
+        data = list(build_training_data())
+        run_training(dec_model, data, 15)
+        torch.save(dec_model.state_dict(), "decoder.pt")
 
 
     with open("predictions.json") as f:
@@ -110,7 +113,7 @@ if __name__ == "__main__":
     acts = torch.FloatTensor(p["sg"])
 
     while len(phones) < 10:
-        next_phone = dec_model(torch.cat(acts, last_phone))
+        next_phone = dec_model(torch.cat((acts, last_phone)).to(device))
         idx = torch.argmax(next_phone)
         phone = all_phones[idx]
 

### 2023-12-11 model.py
batch training

@@ -5,6 +5,7 @@ import numpy as np
 import torch
 import torch.nn as nn
 import torch.optim as optim
+from torch.utils.data import DataLoader
 
 from xsampa import cyrillic_to_xsampa
 from wickel import (
@@ -67,15 +68,15 @@ net_both = Net(output_multiplier=2).to(device)
 
 loss_criterion = nn.L1Loss().to(device)
 
+BATCH_SIZE = 1024
 
 def run_training(model, mapping_pairs, num_epochs):
     optimizer = optim.SGD(model.parameters(), lr=15.0)
 
     for epoch in range(num_epochs):
-        for input_rep, target_rep in mapping_pairs:
-            # TODO: this is copied from the problem set, do we actually need it?
-            input_rep = input_rep.view(1, -1)
-            target_rep = target_rep.view(1, -1)
+        for input_rep, target_rep in DataLoader(mapping_pairs, batch_size=BATCH_SIZE):
+            input_rep = input_rep.to(device)
+            target_rep = target_rep.to(device)
 
             optimizer.zero_grad()
             output = model(input_rep)
@@ -130,9 +131,9 @@ if __name__ == "__main__":
     both_maps = []
 
     for nom_act, sg_act, pl_act in build_activations():
-        sg_maps.append((nom_act.to(device), sg_act.to(device)))
-        pl_maps.append((nom_act.to(device), pl_act.to(device)))
-        both_maps.append((nom_act.to(device), torch.cat((sg_act, pl_act)).to(device)))
+        sg_maps.append((nom_act, sg_act))
+        pl_maps.append((nom_act, pl_act))
+        both_maps.append((nom_act, torch.cat((sg_act, pl_act))))
 
     # TODO: start train/test splitting to check for overfitting
 

### 2023-12-11 new_decoder.py

@@ -1,3 +1,4 @@
+import itertools
 import json
 import os.path
 
@@ -5,6 +6,7 @@ import numpy as np
 import torch
 import torch.nn as nn
 import torch.optim as optim
+from torch.utils.data import DataLoader
 
 from wickel import (
     WickelPhone,
@@ -53,21 +55,19 @@ dec_model = Decoder().to(device)
 
 loss_criterion = nn.L1Loss().to(device)
 
+BATCH_SIZE = 16 * 1024
 
 def run_training(model, mapping_pairs, num_epochs):
     optimizer = optim.SGD(model.parameters(), lr=0.1)
 
     for epoch in range(num_epochs):
-        for i, (input_rep, target_rep) in enumerate(mapping_pairs):
-            input_rep = input_rep.to(device).view(1, -1)
-            target_rep = target_rep.to(device).view(1, -1)
-
+        for i, (input_rep, target_rep) in enumerate(DataLoader(mapping_pairs, batch_size=BATCH_SIZE)):
             optimizer.zero_grad()
-            output = model(input_rep)
-            loss = loss_criterion(output, target_rep)
+            output = model(input_rep.to(device))
+            loss = loss_criterion(output, target_rep.to(device))
             loss.backward()
             optimizer.step()
-            print(f"trained {i + 1}/{len(mapping_pairs)}", end="\r")
+            print(f"trained {((i + 1) * BATCH_SIZE)}/{len(mapping_pairs)}", end="\r")
 
         print(f"loss at epoch {epoch}: {loss.item():.6f}")
 
@@ -99,9 +99,10 @@ if __name__ == "__main__":
         dec_model.load_state_dict(torch.load("decoder.pt"))
         dec_model.eval()
     else:
-        print("beginning training")
+        print("building data")
         data = list(build_training_data())
-        run_training(dec_model, data, 15)
+        print("beginning training")
+        run_training(dec_model, data, 50)
         torch.save(dec_model.state_dict(), "decoder.pt")
 
 

### 2023-12-11 new_decoder.py
lazily generate data

@@ -99,10 +99,8 @@ if __name__ == "__main__":
         dec_model.load_state_dict(torch.load("decoder.pt"))
         dec_model.eval()
     else:
-        print("building data")
-        data = list(build_training_data())
         print("beginning training")
-        run_training(dec_model, data, 50)
+        run_training(dec_model, build_training_data(), 50)
         torch.save(dec_model.state_dict(), "decoder.pt")
 
 
